Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-23 ICML2019/main.tex
tweak

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale
proportional to $\mathcal{X}$ (the dimension of the parameter space) to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

@@

\noindent \textbf{A first approach: the $\expmech$ mechanism.}
\label{sec_ehd}
$\expmech$'s pseudocode is shown in Algorithm \ref{mech:expmech1}. $\expmech$ is an instance
of the Exponential mechanism where the scoring function is instantiated with the function: $r\mapsto -\hellinger(\bysinfer{\dataobs}, r)$.
Given the input $\dataobs\in\mathcal{X}^n$ and a prior $\dirichlet{\valpha}$, the mechanism $\expmech$ outputs an element $r\in\candidateset{\valpha}$ with  probability: 
\[
\frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{2\cdot GS})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{2\cdot GS})},
\]
where $GS$ is the global sensitivity of the scoring function, defined as:
\[
\max_{\dataobs,\dataobs': \\ \adj{\dataobs}{\dataobs'}}\max_{r\in \candidateset{\valpha}}
|\hellinger(\bys{\dataobs}{\valpha}, r) - \hellinger(\bys{\dataobs'}{\valpha}, r)|
\]
% 
  \begin{algorithm}
  \caption{$\expmech$}
  \label{mech:expmech1}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf set} $z=r$ with probability $\frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{2\cdot GS})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{2\cdot GS})}$
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}
\begin{lem}
  Algorithm \ref{mech:expmech1} is $\epsilon$-d.p.
\end{lem}

@@

\begin{thm}
  Algorithm \ref{mech:sexpmech} is $\epsilon$-d.p.
\end{thm}
% As already mentioned $\hexpmech$ extends straightforwardly to the Dirichlet-multinomial system.
% Indeed, the Hellinger distance between two Dirichlet distribution, unfolds in:
% \[
%   \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}},
% \]
% and the $\candidateset$ will consist of the set of all possible posterior Dirichlet
% distributions candidates. Also, the smooth sensitivity $S(\dataobs)$
% in (\ref{thm_gamma_smooth}) will be computed by letting $\dataobs'$ range
% over all the elements in $\datauni^{n}$ adjacent to $\dataobs$. 
\section{Accuracy Analysis}
We will observe accuracy of mechanisms.
In this section, we fix $\alpha$ and $\beta$, and for simplicity, we define the term of difference from $\betad(\alpha, \beta)$, the true posterior distribution.
\[
c(t) = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t).
\]
\subsection{Accuracy Bound for Baseline Mechanisms}

>>>>

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale
proportional to $\mathcal{X}$ (the dimension of the parameter space) to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-differentially private.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-differentially private.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

@@

\noindent \textbf{A first approach: the $\expmech$ mechanism.}
\label{sec_ehd}
$\expmech$'s pseudocode is shown in Algorithm \ref{mech:expmech1}. $\expmech$ is an instance
of the Exponential mechanism where the scoring function is instantiated with the function: $r\mapsto -\hellinger(\bysinfer{\dataobs}, r)$.
Given the input $\dataobs\in\mathcal{X}^n$ and a prior $\dirichlet{\valpha}$, the mechanism $\expmech$ outputs an element $r\in\candidateset{\valpha}$ with  probability: 
\[
\frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{2\cdot GS})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{2\cdot GS})},
\]
where $GS$ is the global sensitivity of the scoring function, defined as:
\[
\max_{\dataobs,\dataobs': \\ \adj{\dataobs}{\dataobs'}}\max_{r\in \candidateset{\valpha}}
|\hellinger(\bys{\dataobs}{\valpha}, r) - \hellinger(\bys{\dataobs'}{\valpha}, r)|
\]
% 
  \begin{algorithm}
  \caption{$\expmech$}
  \label{mech:expmech1}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf set} $z=r$ with probability $\frac
  {\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r)}{2\cdot GS})}
{\Sigma_{r' \in \candidateset{\valpha}}\exp(\frac{-\epsilon\cdot \hellinger(\bys{\dataobs}{\valpha}, r')}{2\cdot GS})}$
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}
\begin{lem}
  Algorithm \ref{mech:expmech1} is $\epsilon$-differentially private.
\end{lem}

@@

\begin{thm}
  Algorithm \ref{mech:sexpmech} is $\epsilon$-differentially private.
\end{thm}
% As already mentioned $\hexpmech$ extends straightforwardly to the Dirichlet-multinomial system.
% Indeed, the Hellinger distance between two Dirichlet distribution, unfolds in:
% \[
%   \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}},
% \]
% and the $\candidateset$ will consist of the set of all possible posterior Dirichlet
% distributions candidates. Also, the smooth sensitivity $S(\dataobs)$
% in (\ref{thm_gamma_smooth}) will be computed by letting $\dataobs'$ range
% over all the elements in $\datauni^{n}$ adjacent to $\dataobs$. 
\section{Accuracy Analysis}
We will observe accuracy of mechanisms.
In this section, we fix $\alpha$ and $\beta$, and for simplicity, we define the term of difference from $\betad(\alpha, \beta)$, the true posterior distribution.
\[
c(t) = \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t).
\]
\subsection{Accuracy Bound for Baseline Mechanisms}
